---
title: "Lab 03 - Nobel laureates"
author: "Heather Hawkins"
date: "01/23/23"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
nobel <- read_csv("data/nobel.csv")
```

## Exercises

### Exercise 1

```{r data, message=FALSE}
nobel <- read_csv("data/nobel.csv")
dim(nobel)
#to specify columns and rows
```

26 variables and 935 observations, each observation is a Noble Laureate

### Exercise 2


```{r nobel_living}
nobel_living <- nobel %>%
  filter(!is.na(country), gender!="org",is.na(died_date),)

dim(nobel_living)
```

Got it! ☺️

```{r USA}
living_during_prize <- nobel_living %>%
  mutate(
    living_during_prize = if_else(country == "USA", "USA", "Other")
  )
```

```{R Sciences}
nobel_living_science <- living_during_prize %>%
  filter(category %in% c("Physics", "Medicine", "Chemistry", "Economics"))

dim(nobel_living)
```

### Exercise 3

```{R prize-us}
nobel_living_science_graph <- nobel_living_science %>%
  group_by(category,living_during_prize) %>%
  summarize(country_count = table(living_during_prize))
ggplot(nobel_living_science_graph, mapping = aes(x= country_count, y = living_during_prize))+
  geom_bar(position = "dodge", stat = "identity")+ facet_wrap(~category)+ labs (x= "Number of Nobel Laureates", y= "Living When Recieving Prize") 

```

### Exercise 4
```{r USA_born}
nobel_living_science_born <- nobel_living_science %>%
  mutate(
    born_country_us = if_else(born_country == "USA", "USA", "Other")
  )

```

```{R prize-born}
ggplot(nobel_living_science_born, aes(x = born_country_us)) +
  geom_bar(position = "stack") +
  coord_flip() + facet_wrap(~category)+ labs (x= "Country of Birth", y= "Number of Nobel Laureates" ) + labs(fill="Category of Science") 

```

```{r nobel_living_NA}
nobel_living_science_born %>%
  group_by(born_country_us) %>% 
  dplyr::count(born_country_us)

```
It seems as if 105 winners were born in the USA
graphBorn_Country <- nobel_living_science_born %>%
  group_by(category,living_during_prize,born_country_us) %>%
  summarize(country_count=table(nobel_living_science_born))

### Exercise 5
```{r USA_born_prize, fig.height=10, fig.width=10}
ggplot(nobel_living_science_born, aes(x = living_during_prize, fill= born_country_us)) +
  geom_bar(position = "stack") +
  coord_flip() + facet_wrap(~category)+ labs (x= "Living During Award", y= "Number of Nobel Laureates" ) + labs(fill="Place of Birth") 

```
Although most U.S. based Laureates were born in the US for the sciences, there is a significant amount of U.S Lauretes that were born in other countries. Because of this, I do believe this data supports Buzzfeed's claim. 
### Exercise 6

...
